### 2020-12-07 Thesis-Chapter3-PolicyMaking.tex
Updates based on presentation feedback
Work on wicked vs tame descriptions

\begin{displayquote}
  \begin{enumerate}
    \item There is no definitive formulation of a wicked problem [and each formulation presupposes a solution].
    \item Wicked problems have no stopping rule.
    \item Solutions to wicked problems are not true-or-false, but good-or-bad.
    \item There is no immediate and no ultimate test of a solution to a wicked problem.
    \item Every solution to a wicked problem is a ``one-shot operation''; because there is no opportunity to learn by
          trial-and-error, every attempt counts significantly.
    \item Wicked problems do not have an enumerable (or an exhaustively describable) set of potential solutions, nor is
          there a well-described set of permissible operations that may be incorporated into the plan.
    \item Every wicked problem is essentially unique.
    \item Every wicked problem can be considered to be a symptom of another problem.
    \item The existence of a discrepancy representing a wicked problem can be explained in numerous ways. The choice of
          explanation determines the nature of the problem's resolution.
    \item The planner has no right to be wrong [i.e., they are liable for the consequences of their decisions].
  \end{enumerate}
\end{displayquote}

Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
conditions.

Several of these characteristics are results of the fact that we have no accurate predictive model of the world and
human behavior at large. This does not need to be expounded upon.

The more important insight is one step removed: precisely because there is an inexhaustible set of potential solutions,
the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
proposed solution demands research and context to be added to the problem definition. This feedback from proposal to
definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals to
a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
themselves.

@@

Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
amount will result in a certain difference in response with a certain statistical confidence. Refining the problem
statement and combining the tame elements in a cohesive solution is the ``wicked'' part. Despite certain tame elements,
one cannot verify the effectiveness of a plan ahead of time, port a solution from a different domain, or even
definitively evaluate its performance after the fact.

@@

@@

\section{Proposal: The OODA Loop for Wicked Problems}

@@

>>>>

\begin{displayquote}
  \begin{enumerate}
    \item There is no definitive formulation of a wicked problem [and each formulation presupposes a solution].
    \item Wicked problems have no stopping rule.
    \item Solutions to wicked problems are not true-or-false, but good-or-bad.
    \item There is no immediate and no ultimate test of a solution to a wicked problem.
    \item Every solution to a wicked problem is a ``one-shot operation''; because there is no opportunity to learn by
          trial-and-error, every attempt counts significantly.
    \item Wicked problems do not have an enumerable (or an exhaustively describable) set of potential solutions, nor is
          there a well-described set of permissible operations that may be incorporated into the plan.
    \item Every wicked problem is essentially unique.
    \item Every wicked problem can be considered to be a symptom of another problem.
    \item The existence of a discrepancy representing a wicked problem can be explained in numerous ways. The choice of
          explanation determines the nature of the problem's resolution.
    \item The [decision maker] has no right to be wrong [i.e., they are liable for the consequences of their decisions].
  \end{enumerate}
\end{displayquote}

Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
conditions. The ``wicked'' problem's counterpart is the ``tame'' problem. ``Tame'' does not imply easy; it only means
that the scientific approach will be effective.

Several of \acp{wicked-prob}' characteristics are results of the fact that we have no accurate predictive model of the
world and human behavior at large. This does not need to be expounded upon.

The more important insight is one step removed: precisely because there is an inexhaustible set of potential solutions,
the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
proposed solution demands research and context to be fed back into the problem definition. This feedback from proposal
to definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals
to a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
themselves.

@@

\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
solution in a way that doesn't leave vast swaths of people behind.

% TODO: what am I really trying to express here?
% ... WPs quickly grow in complexity with the number of subproblems that comprise them. Refining the problem statement
% and combining tame solutions with wicked subproblem mitigations is the truly ``wicked'' part.

% Elements of the problem may be tame---for example, one can independently confirm that reducing global carbon emissions
% by a certain amount will result in a certain difference in response with a certain statistical confidence. Other
% elements of the problem are wicked---for example, one cannot verify the effectiveness of a plan ahead of time, port a
% solution from a different domain, or definitively evaluate its performance after the fact.

% Refining the problem statement and combining the
% tame elements into a cohesive solution is the ``wicked'' part.

@@

% Prof. Engelsma
% - A big part of the "solution" is getting people together
% - My thesis seems aimed at tech people to get them to understand and judge threats
% - How can I take this accessible and bridge the gap for policymakers?

@@

% TODO: Dulimarta: consider changing this heading to "Proposal: Observe-Orient-Decide-Act"
\section{Proposal: The OODA Loop for Wicked Problems}

@@

% TODO: explain the icons
